### 2025-09-14 paper/sec/introduction.tex
High-resolution screen capture. Fix filename to match example.

\begin{itemize}
    \item Interactive visualization of data
    \item Making the text transparent, with the possibility to highlight the data related to a specific sentence.
    \item Making sense of data-driven claims is hard, even if the evidence base (code/data) is open
    \item We see this in peer review, misinformation, retracted papers, …
\end{itemize}

\subsection{Transparent Text}

Recent work on data provenance and data visualisation has shown how interactive provenance queries can be
supported automatically \cite{psallisas

\begin{itemize}
    \item Introduce basic idea, as complementary to transparent visualisations
\end{itemize}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{fig/scigen-mr-1906.02780-10.png}
    \caption{Sample text with interpretation generated by Authoring Assistant}\label{fig:scigen-example-website}
\end{figure}

>>>>

\begin{itemize}
    \item Making the text transparent, with the possibility to highlight the data related to a specific sentence.
    \item Making sense of data-driven claims is hard, even if the evidence base (code/data) is open
    \item We see this in peer review, misinformation, retracted papers, …
\end{itemize}

\subsection{Transparent Text via Interactive Provenance Queries}

Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} integrating
interactive provenance queries into charts and figures using \emph{transparent programming languages} which
track data provenance automatically as data flows through a computation. Visual outputs produced using these
platforms allow readers to interact with visual elements to see the relevant data. In most article charts and
figures usually only play a supporting role, however. The main claims are usually presented in natural
language which also makes quantitative claims and refers to data analyses which are not present in the final
artifact.

In principle transparent programming languages can also be used to create natural language which is bidirectly
linked to data, by computing key fragments of natural language from data. For example, using the language
presented in this paper,

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-with-pointer.png}
    \caption{Transparent Text with Interactive Provenance Queries}\label{fig:scigen-example-website}
\end{figure}

\rpnote{Small example here}

@@
